\documentclass[conference]{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{newtxtext,newtxmath}
\usepackage{amsmath,bm}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{cite}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[breaklinks=true,hidelinks]{hyperref}

% ---- Code listing style -----------------------------------------------
\lstdefinestyle{cppstyle}{
    language=C++,
    basicstyle=\ttfamily\scriptsize,
    keywordstyle=\color{blue!70!black}\bfseries,
    commentstyle=\color{gray}\itshape,
    stringstyle=\color{orange!80!black},
    breaklines=true,
    breakatwhitespace=true,
    frame=single,
    rulecolor=\color{black!30},
    xleftmargin=4pt,
    xrightmargin=4pt,
    numbers=none,
    tabsize=2,
}
\lstset{style=cppstyle}

% -----------------------------------------------------------------------

\title{Embedded-Friendly IMU Calibration:\\
Ellipsoid Fitting, Temperature Compensation,\\
and Robust Estimation in C\texttt{++}/Eigen}

\author{%
    \IEEEauthorblockN{Mikhail S.\ Grushinskiy}
    \IEEEauthorblockA{Independent Researcher\\
    Bareboat Necessities Project\\
    \texttt{mgrouch@users.github.com}}
}

\begin{document}
    \maketitle

% -----------------------------------------------------------------------
    \begin{abstract}
        This note documents \texttt{imu\_cal}, a modern, header-only C\texttt{++}
        library for in-situ calibration of the three sensor axes of an Inertial
        Measurement Unit (IMU): accelerometer, magnetometer, and gyroscope.
        The library is designed for Arduino-class embedded targets using the
        Eigen linear-algebra library.
        It fits each sensor to an ellipsoidal model via a robust algebraic
        10-parameter least-squares solver, applies trimmed iterative reweighting
        to suppress outliers, enforces positive-definiteness of the recovered
        shape matrix, and optionally compensates for temperature-dependent
        bias drift.
        Accelerometer calibration additionally enforces axis alignment and
        gravity renormalization.
        Magnetometer calibration works in a numerically scaled coordinate
        frame to maintain well-conditioned matrices at microtesla magnitudes.
        All calibrators require a minimum of 80 samples and enforce
        geometric coverage checks before fitting.
        The public API is a small collection of typed structs and two-line
        \texttt{fit()} calls; the complete implementation is contained
        in a single \texttt{.h} file of under 900 lines.
    \end{abstract}

    \begin{IEEEkeywords}
        IMU calibration, ellipsoid fitting, robust estimation,
        temperature compensation, Eigen, embedded systems, Arduino,
        accelerometer, magnetometer, gyroscope
    \end{IEEEkeywords}

% -----------------------------------------------------------------------
    \section{Introduction}

    Low-cost MEMS IMUs suffer from well-known systematic errors:
    scale-factor mismatches, inter-axis cross-sensitivity, constant bias,
    and temperature-dependent bias drift \cite{Frosio2009}.
    Without calibration, these errors accumulate rapidly in attitude
    and heading estimates.

    Classical batch calibration techniques require a precision turntable
    or a known-orientation jig.
    In contrast, \emph{in-situ} or \emph{motional} calibration methods
    collect raw sensor data while the device is rotated freely and
    then recover the error model from the geometry of the measurements
    \cite{Vasconcelos2011}.
    For the accelerometer and magnetometer, the underlying observation
    is that the true sensor output should lie on a sphere of known
    radius ($g$ for acceleration, the local field magnitude $B$ for
    the magnetometer).
    Systematic errors deform this sphere into an ellipsoid; recovering
    the ellipsoid and inverting the distortion yields the calibration map.

    \texttt{imu\_cal} implements this approach for all three sensor axes
    in a self-contained, dependency-light C\texttt{++} header suitable
    for deployment on Arduino-class hardware.
    The library uses Eigen \cite{Eigen} for all matrix operations
    and exposes a minimal, typed API.

% -----------------------------------------------------------------------
    \section{Sensor Error Models}

    \subsection{Accelerometer and Magnetometer}

    Let $\mathbf{v}_\text{raw} \in \mathbb{R}^3$ be the raw sensor reading.
    The standard affine error model is
    \begin{equation}
        \mathbf{v}_\text{raw} = \mathbf{M}\,\mathbf{v}_\text{true} + \mathbf{b}
        + \boldsymbol{\eta},
    \end{equation}
    where $\mathbf{M} \in \mathbb{R}^{3\times 3}$ captures scale factors and
    cross-sensitivity, $\mathbf{b} \in \mathbb{R}^3$ is the constant bias, and
    $\boldsymbol{\eta}$ is noise.
    The calibrated output is therefore
    \begin{equation}
        \mathbf{v}_\text{cal} = \mathbf{S}\bigl(\mathbf{v}_\text{raw} -
        \mathbf{b}(T)\bigr),
        \label{eq:cal_model}
    \end{equation}
    where $\mathbf{S} = \mathbf{M}^{-1}$ is the whitening matrix and
    $\mathbf{b}(T)$ is the temperature-dependent bias (Section~\ref{sec:tempbias}).

    \subsection{Gyroscope}

    The gyroscope error model used here is the simpler bias-only form,
    because scale-factor and cross-sensitivity calibration requires a
    precision rate rig:
    \begin{equation}
        \boldsymbol{\omega}_\text{cal} = \mathbf{S}\bigl(\boldsymbol{\omega}_\text{raw}
        - \mathbf{b}(T)\bigr),
        \quad \mathbf{S} = \mathbf{I}_3.
    \end{equation}
    The identity scale matrix is retained in the struct so that a measured-rate
    rig can upgrade it without changing the API.

    \subsection{Temperature Bias Model}
    \label{sec:tempbias}

    Gyroscope and accelerometer biases are strongly temperature-dependent.
    The library models this as a first-order linear expansion about a
    reference temperature $T_0$ (\SI{25}{\degreeCelsius} by default):
    \begin{equation}
        \mathbf{b}(T) = \mathbf{b}_0 + \mathbf{k}(T - T_0),
        \label{eq:tempbias}
    \end{equation}
    with intercept $\mathbf{b}_0 \in \mathbb{R}^3$ and slope
    $\mathbf{k} \in \mathbb{R}^3$ estimated by least squares from
    per-temperature-bin calibration centers (Section~\ref{sec:accel_fit}).

% -----------------------------------------------------------------------
    \section{Ellipsoid Fitting}
    \label{sec:ellipsoid}

    \subsection{Algebraic Form}

    A general ellipsoid centred at $\mathbf{b}$ satisfying
    $\|\mathbf{A}(\mathbf{x}-\mathbf{b})\| = R$ can be written in the
    quadratic algebraic form
    \begin{equation}
        \mathbf{x}^\top \mathbf{Q}\,\mathbf{x}
        + 2\mathbf{q}^\top\mathbf{x} + c = 0,
    \end{equation}
    where $\mathbf{Q} \in \mathbb{R}^{3\times 3}$ is symmetric positive
    definite, $\mathbf{q} \in \mathbb{R}^3$, and $c \in \mathbb{R}$.
    Expanding the defining equation:
    \begin{align}
        \mathbf{Q} &= \mathbf{A}^\top\mathbf{A}, \\
        \mathbf{q} &= -\mathbf{Q}\mathbf{b}, \\
        c &= \mathbf{b}^\top\mathbf{Q}\mathbf{b} - R^2.
    \end{align}

    \subsection{10-Parameter Regression}
    \label{sec:regression}

    Define the feature vector for sample $\mathbf{x}_i = (X,Y,Z)^\top$:
    \begin{equation}
        \mathbf{d}_i = \bigl[X^2,\,Y^2,\,Z^2,\,2XY,\,2XZ,\,2YZ,\,
        2X,\,2Y,\,2Z,\,1\bigr]^\top.
    \end{equation}
    The algebraic constraint becomes $\mathbf{d}_i^\top\mathbf{p} = 0$,
    where $\mathbf{p} \in \mathbb{R}^{10}$ encodes $\mathbf{Q}$, $\mathbf{q}$, $c$.
    Because the constraint is homogeneous, we fix the normalisation by
    setting the right-hand side to 1 (corresponding to $c = 1$), leading
    to the linear system
    \begin{equation}
        \mathbf{H}\,\mathbf{p} = \mathbf{g},
        \quad
        \mathbf{H} = \sum_i \mathbf{d}_i\mathbf{d}_i^\top,
        \quad
        \mathbf{g} = \sum_i \mathbf{d}_i.
        \label{eq:lsq}
    \end{equation}
    A small Tikhonov ridge
    $\lambda = \lambda_\text{rel}\!\left(\tfrac{\mathrm{tr}(\mathbf{H})}{10}
    + 10^{-12}\right)$
    is added to the diagonal of $\mathbf{H}$ before solving via LDLT
    decomposition:
    \begin{equation}
    (\mathbf{H} + \lambda\mathbf{I})\,\mathbf{p} = \mathbf{g}.
    \label{eq:ridge}
    \end{equation}
    The default ridge scale is $\lambda_\text{rel} = 10^{-6}$, which is
    harmless for normal data but prevents numerical failure with
    near-planar samples.

    \subsection{Model Reconstruction}

    From the solution $\mathbf{p}$, the centre and shape matrix are recovered as:
    \begin{align}
        \mathbf{b} &= -\mathbf{Q}^{-1}\mathbf{q}, \\
        s &= 1 + \mathbf{b}^\top\mathbf{Q}\mathbf{b} - c, \\
        \mathbf{M} &= \mathbf{Q}/s.
    \end{align}
    The matrix $\mathbf{M}$ is projected to symmetric positive definite (SPD)
    via eigenvalue clamping (Section~\ref{sec:spd}), then Cholesky-factored
    $\mathbf{M} = \mathbf{U}^\top\mathbf{U}$ to obtain the unit-sphere
    whitening matrix $\mathbf{A}_\text{unit} = \mathbf{U}$.
    The final calibration matrix scaled to the target radius $R$ is
    \begin{equation}
        \mathbf{A} = \mathbf{A}_\text{unit}\,R.
    \end{equation}

% -----------------------------------------------------------------------
    \section{Robust Fitting}
    \label{sec:robust}

    A single pass of algebraic fitting is sensitive to outliers, animal
    interference near the magnetometer, and the unavoidable non-uniformity
    of free-motion data collection.
    The library therefore wraps the linear solve inside an iterative
    trimming loop.

    \subsection{Algorithm Overview}

    \begin{enumerate}
        \item Initialise all sample weights as inliers.
        \item Solve \eqref{eq:lsq}--\eqref{eq:ridge} on the current inlier set.
        \item Compute residuals $e_i = \|\mathbf{A}(\mathbf{x}_i - \mathbf{b})\| - R$.
        \item Compute the Median Absolute Deviation (MAD) of $\{e_i\}$.
        \item Mark as outliers any sample with $|e_i| > \tau$, where
        $\tau = \min(\tau_\text{MAD},\,\tau_\text{quantile})$.
        \item Repeat from step 2 for a fixed number of iterations
        (default 3).
        \item Final solve on the converged inlier set.
    \end{enumerate}

    \subsection{MAD Threshold}

    The MAD-based threshold is
    \begin{equation}
        \tau_\text{MAD} = 5.2\,\max(\mathrm{MAD},\,10^{-4}) + 10^{-6}.
    \end{equation}
    The factor 5.2 is chosen empirically to be permissive enough for
    real sensor noise while still rejecting gross outliers.

    \subsection{Quantile Threshold}

    A trimming fraction $\alpha$ (default \SI{15}{\percent}) is applied by
    sorting the absolute residuals and setting
    \begin{equation}
        \tau_\text{quantile} = |e|_{(\lceil(1-\alpha)n\rceil)},
    \end{equation}
    the $(1-\alpha)$-quantile of the absolute residual distribution.
    Samples above this threshold are excluded.

    \subsection{Fallback}

    If the joint threshold would leave fewer than 10 inliers, the code
    relaxes to the looser of the two thresholds.
    If that still leaves fewer than 10, fitting fails with
    \texttt{INLIERS\_TOO\_FEW}.

% -----------------------------------------------------------------------
    \section{SPD Projection}
    \label{sec:spd}

    Numerical noise can cause the recovered matrix $\mathbf{M}$ to have
    small negative eigenvalues, which would break the subsequent Cholesky
    factorisation.
    The library projects any symmetric matrix to the nearest SPD matrix
    by clamping its eigenvalues:
    \begin{equation}
        \mathbf{M} \leftarrow \mathbf{V}\,
        \mathrm{diag}\!\left(\max(|\lambda_i|,\,\epsilon)\right)\,
        \mathbf{V}^\top,
    \end{equation}
    where $\mathbf{V}$ are the eigenvectors of $\mathbf{M}$ and
    \begin{equation}
        \epsilon = \max\!\bigl(\epsilon_\text{rel}\,\lambda_\text{max},\;
        \epsilon_\text{abs}\bigr),
        \quad
        \epsilon_\text{rel} = 10^{-5},\;
        \epsilon_\text{abs} = 10^{-7}.
    \end{equation}
    This is the nearest-SPD projection under the Frobenius norm for
    symmetric matrices \cite{Higham1988}.

% -----------------------------------------------------------------------
    \section{Coverage and Degeneracy Checks}
    \label{sec:coverage}

    Before fitting, the library verifies that the collected samples
    adequately cover the sphere.
    Let $\hat{\mathbf{x}}_i = (\mathbf{x}_i - \boldsymbol{\mu})/
    \|\mathbf{x}_i - \boldsymbol{\mu}\|$ be the unit-direction vectors
    after mean-centering.

    Three conditions must hold simultaneously:
    \begin{enumerate}
        \item \textbf{Axis span}: the centred sample range on each axis
        must satisfy $\Delta_j \ge 0.90\,R_\text{expected}$.
        \item \textbf{Sign balance}: at least \SI{8}{\percent} of samples
        must fall on each side of zero on every axis.
        \item \textbf{Directional spread}: the direction covariance
        $\mathbf{C} = \frac{1}{m}\sum_i \hat{\mathbf{x}}_i
        \hat{\mathbf{x}}_i^\top$
        must satisfy $\det(\mathbf{C}) \ge 10^{-3}$.
        For a uniform sphere $\mathbf{C} \approx \tfrac{1}{3}\mathbf{I}$
        with $\det \approx 0.037$; the threshold rejects planar or
        linear sample distributions.
    \end{enumerate}
    Failure in any condition produces \texttt{DEGENERATE\_COVERAGE}.

% -----------------------------------------------------------------------
    \section{Accelerometer Calibration}
    \label{sec:accel_fit}

    \subsection{Static Sample Gating}

    Only samples collected while the IMU is near-stationary are useful for
    accelerometer calibration.
    The \texttt{addSample} method enforces two gates:
    \begin{align}
        \bigl|\|\mathbf{a}_\text{raw}\| - g\bigr| &< \delta_a
        \quad (\delta_a = \SI{0.25}{\meter\per\second\squared}), \\
        \|\boldsymbol{\omega}_\text{raw}\| &< \delta_\omega
        \quad (\delta_\omega = \SI{0.07}{\radian\per\second}).
    \end{align}
    Samples not passing both gates are silently discarded.

    \subsection{Temperature Binning}

    Samples are partitioned into $K = 8$ temperature bins of equal width
    spanning $[T_\text{min}, T_\text{max}]$.
    Each bin with at least 30 samples is independently fitted by
    \texttt{ellipsoid\_to\_sphere\_robust} with target radius $g$.

    \subsection{Axis-Rotation Prevention}
    \label{sec:polar}

    The raw whitening matrix $\mathbf{A}$ returned by the ellipsoid fit
    may contain a rigid rotation component that would misalign the calibrated
    axes.
    To remove it, the library extracts the \emph{polar SPD factor}:
    \begin{equation}
        \mathbf{P} = \sqrt{\mathbf{A}^\top\mathbf{A}},
        \quad\text{so that}\quad
        \mathbf{A} = \mathbf{R}\,\mathbf{P},\;
        \mathbf{P} \succ 0.
    \end{equation}
    $\mathbf{P}$ captures scale and cross-sensitivity while discarding any
    rotation $\mathbf{R}$.
    Optionally, only the diagonal of $\mathbf{P}$ is retained
    (\texttt{AccelSMode::DiagonalOnly}, the default), which further
    prevents axis permutation at the cost of ignoring minor cross-axis
    terms.

    \subsection{Plausibility Gates}

    A bin's $\mathbf{S}$ matrix is accepted only if:
    \begin{itemize}
        \item all diagonal entries lie in $[0.80,\,1.25]$,
        \item the SPD condition number satisfies $\kappa(\mathbf{S}) \le 3.5$,
        \item in \texttt{PolarSPD} mode, the off-diagonal RMS satisfies
        $\text{rms}_\text{off}(\mathbf{S}) \le 0.05$.
    \end{itemize}
    These thresholds reject numerically or physically implausible fits
    (\texttt{ACCEL\_S\_UNPHYSICAL}).

    \subsection{Gravity Renormalisation}

    After fitting, the scalar magnitude of the calibrated output is
    matched to $g$ using a robust median of quiet-period norms:
    \begin{equation}
        \mathbf{S} \leftarrow \frac{g}{\mathrm{med}\bigl(
        \|\mathbf{S}(\mathbf{a}_i - \mathbf{b}(T_i))\|\bigr)}\,\mathbf{S}.
    \end{equation}
    The scale factor is clamped to $[0.97,\,1.03]$ (relaxing to
    $[0.90,\,1.10]$ on fallback).
    This step corrects for any remaining uniform scale error without
    rotating the axes.

% -----------------------------------------------------------------------
    \section{Magnetometer Calibration}
    \label{sec:mag_fit}

    \subsection{Numeric Scaling}

    Raw magnetometer values are typically in the range
    $\sim\SI{25}{}$--$\SI{65}{\micro\tesla}$, which leads to condition
    numbers of order $10^6$ in the Gram matrix $\mathbf{H}$
    of \eqref{eq:lsq}.
    The library therefore normalises all samples by a robust radius estimate:
    \begin{equation}
        \mathbf{x}' = \mathbf{x} / \hat{B},
        \quad
        \hat{B} = \mathrm{med}\bigl(\|\mathbf{x}_i - \boldsymbol{\mu}\|\bigr),
    \end{equation}
    fits the unit-sphere model ($R'=1$) in the normalised frame, then
    unscales the result:
    \begin{align}
        \mathbf{b}_{\si{\micro\tesla}} &= \hat{B}\,\mathbf{b}', \\
        \mathbf{A}_{\si{\micro\tesla}} &= \frac{\mathbf{A}'}{{\hat{B}}}\,B_\text{med},
    \end{align}
    where $B_\text{med}$ is the median corrected radius in physical units.

    \subsection{Field Plausibility Check}

    The recovered field magnitude $B_\text{med}$ is accepted only if
    $12 \le B_\text{med} \le \SI{120}{\micro\tesla}$.
    This rejects degenerate solutions and catches situations where
    the sensor was near a strong magnetic disturbance during the
    entire collection.

% -----------------------------------------------------------------------
    \section{Gyroscope Calibration}
    \label{sec:gyro_fit}

    The gyroscope calibrator collects stationary samples gated by:
    \begin{align}
        \|\boldsymbol{\omega}_\text{raw}\| &< \SI{0.07}{\radian\per\second}, \\
        \bigl|\|\mathbf{a}_\text{raw}\| - g\bigr| &< \SI{0.25}{\meter\per\second\squared}.
    \end{align}

    Samples are binned by temperature (same $K=8$ scheme as the accelerometer).
    Within each bin, the mean reading is taken as the bias estimate.
    If at least two bins are populated, a linear temperature model
    \eqref{eq:tempbias} is fitted by ordinary least squares.
    Otherwise a constant bias ($\mathbf{k}=\mathbf{0}$) is used.

% -----------------------------------------------------------------------
    \section{Software Architecture}
    \label{sec:arch}

    \subsection{Namespace and Configuration}

    All symbols are enclosed in namespace \texttt{imu\_cal}.
    The compile-time constant
    \begin{lstlisting}
static constexpr int IMU_CAL_MAX_SAMPLES = 400;
    \end{lstlisting}
    limits the size of all fixed-capacity arrays to prevent dynamic
    allocation.

    \subsection{Error Reporting}

    Failures are reported through a typed enum:
    \begin{lstlisting}
enum class FitFail : uint8_t {
  OK = 0,
  BAD_ARG, TOO_FEW_SAMPLES,
  NON_FINITE_INPUT,
  DEGENERATE_COVERAGE,
  SOLVE_LDLT_FAIL,
  MODEL_LU_NOT_INVERTIBLE,
  MODEL_S_NONPOSITIVE,
  MODEL_SPD_PROJECT_FAIL,
  MODEL_CHOLESKY_FAIL,
  INLIERS_TOO_FEW,
  MAG_FIELD_OUT_OF_RANGE,
  TEMP_BINS_EMPTY,
  ACCEL_S_UNPHYSICAL,
};
    \end{lstlisting}
    The helper \texttt{fitFailStr()} converts any code to a printable
    C string for serial diagnostics.

    \subsection{Sample Buffer}

    \texttt{SampleBuffer3<T,N>} is a fixed-capacity ring buffer for
    3-D vectors with associated temperature readings:
    \begin{lstlisting}
template <typename T, int N>
struct SampleBuffer3 {
  Eigen::Matrix<T,3,1> v[N];
  T    tempC[N];
  int  n = 0;
  bool push(const Vec3& x, T tC = T(0));
};
    \end{lstlisting}
    \texttt{push} validates finiteness and rejects samples beyond capacity.

    \subsection{Calibration Result Structs}

    Each sensor has a corresponding result struct:
    \begin{itemize}
        \item \texttt{AccelCalibration<T>}: stores $g$, $\mathbf{S}$,
        \texttt{TempBias3}, RMS, and an \texttt{apply()} method.
        \item \texttt{MagCalibration<T>}: stores $\mathbf{A}$, $\mathbf{b}$,
        $B_\text{field}$, RMS, and an \texttt{apply()} method.
        \item \texttt{GyroCalibration<T>}: stores $\mathbf{S}$,
        \texttt{TempBias3}, and an \texttt{apply()} method.
    \end{itemize}
    All structs carry a Boolean \texttt{ok} flag and use
    \texttt{EIGEN\_MAKE\_ALIGNED\_OPERATOR\_NEW} for correct Eigen alignment.

    \subsection{Calibrator Templates}

    Three template calibrators own a sample buffer and expose a
    two-method API: \texttt{addSample} and \texttt{fit}.

    \begin{lstlisting}
template <typename T, int N, int K_TBINS = 8>
struct AccelCalibrator {
  bool addSample(const Vec3& a, const Vec3& w, T tempC);
  bool fit(AccelCalibration<T>& out, ...);
};

template <typename T, int N>
struct MagCalibrator {
  bool addSample(const Vec3& m_uT);
  bool fit(MagCalibration<T>& out, ...);
};

template <typename T, int N, int K_TBINS = 8>
struct GyroCalibrator {
  bool addSample(const Vec3& w, const Vec3& a, T tempC);
  bool fit(GyroCalibration<T>& out, ...);
};
    \end{lstlisting}

% -----------------------------------------------------------------------
    \section{Typical Usage}
    \label{sec:usage}

    The following snippet illustrates a typical calibration workflow
    for a float-precision magnetometer with 200 sample capacity:

    \begin{lstlisting}
#include "imu_cal.h"
using namespace imu_cal;

MagCalibrator<float, 200> cal;

// -- collection phase (rotate device freely) --
while (!cal.buf.n < 100) {
  Eigen::Vector3f m_raw;
  readMag(m_raw);          // application-specific
  cal.addSample(m_raw);
}

// -- fitting phase --
MagCalibration<float> result;
FitFail why;
if (!cal.fit(result, 3, 0.15f, 1e-6f, &why)) {
  Serial.println(fitFailStr(why));
  return;
}

// -- application phase --
Eigen::Vector3f m_cal = result.apply(m_raw);
    \end{lstlisting}

    The accelerometer and gyroscope calibrators additionally require a
    simultaneous gyroscope / accelerometer reading for their respective
    static-motion gates.

% -----------------------------------------------------------------------
    \section{Numerical Notes}
    \label{sec:numerical}

    \subsection{Precision}

    The library is parameterised by scalar type \texttt{T}.
    \texttt{float} (32-bit) is sufficient for field deployment.
    \texttt{double} (64-bit) is recommended when validating the algorithm
    on a host machine.
    Internally, all \texttt{sqrt}, \texttt{fabs}, and \texttt{isfinite}
    calls are cast through \texttt{double} to avoid platform-specific
    float precision issues on AVR-class targets.

    \subsection{No Dynamic Allocation}

    All internal arrays (residuals, inlier mask, sorted quantile buffer)
    are declared at fixed size \texttt{IMU\_CAL\_MAX\_SAMPLES}.
    The AccelCalibrator keeps its per-bin index arrays in a mutable
    \texttt{Workspace} member to reduce stack pressure inside \texttt{fit()}.

    \subsection{Eigen Dependency}

    The header conditionally includes either \texttt{<Eigen/Dense>} and
    \texttt{<Eigen/Eigenvalues>} (host builds) or
    \texttt{<ArduinoEigenDense.h>} (Arduino builds):
    \begin{lstlisting}
#ifdef EIGEN_NON_ARDUINO
#  include <Eigen/Dense>
#  include <Eigen/Eigenvalues>
#else
#  include <ArduinoEigenDense.h>
#endif
    \end{lstlisting}
    No other external dependencies are required.

% -----------------------------------------------------------------------
    \section{Failure Mode Summary}

    Table~\ref{tab:failures} summarises the failure codes, their root
    causes, and typical remedies.

    \begin{table}[t]
        \centering
        \caption{Failure codes and remedies}
        \label{tab:failures}
        \renewcommand{\arraystretch}{1.2}
        \begin{tabular}{@{}lp{4.2cm}@{}}
            \toprule
            \textbf{Code} & \textbf{Meaning / Remedy} \\
            \midrule
            \texttt{TOO\_FEW\_SAMPLES}      & Collect at least 80 samples per bin \\
            \texttt{NON\_FINITE\_INPUT}     & Check ADC / I$^2$C link for errors \\
            \texttt{DEGENERATE\_COVERAGE}   & Rotate more; cover all six hemisphere faces \\
            \texttt{SOLVE\_LDLT\_FAIL}      & Increase ridge; unlikely with normal data \\
            \texttt{MODEL\_S\_NONPOSITIVE}  & Sensor may be saturated; check range \\
            \texttt{MODEL\_SPD\_PROJECT\_FAIL} & Near-degenerate geometry; collect more data \\
            \texttt{MODEL\_CHOLESKY\_FAIL}  & SPD projection failed; rarely occurs \\
            \texttt{INLIERS\_TOO\_FEW}      & Too many outliers; check mounting, vibration \\
            \texttt{MAG\_FIELD\_OUT\_OF\_RANGE} & Keep away from strong magnets; recheck sensor \\
            \texttt{TEMP\_BINS\_EMPTY}      & Too few samples per temperature bin; slow down cooling \\
            \texttt{ACCEL\_S\_UNPHYSICAL}   & Diagonal or condition check failed; check gates \\
            \bottomrule
        \end{tabular}
    \end{table}

% -----------------------------------------------------------------------
    \section{Conclusion}

    \texttt{imu\_cal} provides a complete, self-contained in-situ IMU
    calibration pipeline in fewer than 900 lines of C\texttt{++}/Eigen.
    The core algorithm is a 10-parameter ellipsoid fit solved by LDLT
    with iterative trimmed reweighting and MAD-based robust scaling.
    Accelerometer calibration adds temperature binning, polar-SPD
    axis-alignment enforcement, and gravity renormalisation.
    Magnetometer calibration applies a coordinate scaling step to improve
    conditioning at microtesla magnitudes.
    Gyroscope calibration extracts a linear temperature-bias model from
    quiet-stationary data.
    The library is header-only, allocation-free, and configurable via
    template parameters, making it directly deployable on
    Arduino-compatible embedded targets.

% -----------------------------------------------------------------------
    \begin{thebibliography}{9}

        \bibitem{Frosio2009}
        I.~Frosio, F.~Pedersini, and N.~A.~Borghese,
        ``Autocalibration of MEMS Accelerometers,''
        \emph{IEEE Trans.\ Instrumentation and Measurement},
        vol.~58, no.~6, pp.~2034--2041, Jun.\ 2009.

        \bibitem{Vasconcelos2011}
        J.~F.~Vasconcelos, G.~Elkaim, C.~Silvestre, P.~Oliveira, and B.~Cardeira,
        ``Geometric Approach to Strapdown Magnetometer Calibration in Sensor Frame,''
        \emph{IEEE Trans.\ Aerospace and Electronic Systems},
        vol.~47, no.~2, pp.~1293--1306, Apr.\ 2011.

        \bibitem{Eigen}
        G.~Guennebaud, B.~Jacob, et al.,
        ``Eigen v3,''
        \url{http://eigen.tuxfamily.org}, 2010.

        \bibitem{Higham1988}
        N.~J.~Higham,
        ``Computing a Nearest Symmetric Positive Semidefinite Matrix,''
        \emph{Linear Algebra and its Applications},
        vol.~103, pp.~103--118, 1988.

    \end{thebibliography}

\end{document}

