\documentclass[conference]{IEEEtran}

\usepackage{iftex}
\ifPDFTeX
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\fi
\usepackage{amsthm}
\usepackage{amsmath, bm}
\usepackage{newtxtext,newtxmath}
\usepackage{textcomp}
\interdisplaylinepenalty=2500
\usepackage{pgf}
\usepackage{pgfplots}
\usepackage{xcolor}
\usepackage{float}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{svg}
\usepackage{placeins}
\usepackage{array, booktabs, makecell, adjustbox}
\usepackage{mathtools}
\usepackage{nccmath, relsize}
\usepackage{enumitem}
\usepackage{balance}
\usepackage{flafter}
\usepackage{siunitx}
\usepackage[final]{microtype}
\usepackage{etoolbox}

\DeclareRobustCommand{\vct}[1]{\bm{#1}}
\DeclareRobustCommand{\mat}[1]{\bm{#1}}
\DeclareRobustCommand{\matg}[1]{\bm{#1}}
\newcommand{\Quat}{\mathbb{H}}
\newcommand{\Sph}{\mathbb{S}}
\newcommand{\SO}{\mathrm{SO}}
\newcommand{\SE}{\mathrm{SE}}
\newcommand{\quat}[1]{\vct{#1}}
\newcommand{\Rot}{\bm{\mathcal{R}}}
\newcommand{\qconj}[1]{\overline{#1}}
\newcommand{\qreal}[1]{\operatorname{Re}\! \left(#1\right)}
\newcommand{\qimag}[1]{\operatorname{Im}\! \left(#1\right)}

\usepackage{cite}
\usepackage[breaklinks=true,hidelinks]{hyperref}
\usepackage{bookmark}
\usepackage[nameinlink,capitalise,noabbrev]{cleveref}

\sisetup{detect-weight=true,detect-family=true,per-mode=symbol}
\setlist[description]{font=\footnotesize\bfseries,labelsep=0.6em,leftmargin=1.8em}

\pdfstringdefDisableCommands{%
    \def\SI#1#2{#1 #2}%
    \def\bm#1{#1}%
    \def\Skew#1{[#1]_\text{x}}%
    \def\Id{I}%
    \def\pct#1{#1\%}%
    \def\circ{deg}%
    \def\degree{deg}%
    \def\propto{∝}%
    \def\sigma{sigma}%
    \def\tau{tau}%
    \def\Phi{Phi}%
    \def\omega{omega}%
    \def\_{}%
    \def\cdot{}%
    \def\;{ }%
    \def\I{I}%
    \def\quat#1{#1}%
    \def\SO{SO}%
    \def\SE{SE}%
    \def\Quat{H}%
    \def\Sph{S}%
}

\AtBeginEnvironment{equation}{\small}
\AtBeginEnvironment{align}{\small}
\providecommand{\mathdefault}[1]{#1}

\setcounter{topnumber}{5}
\setcounter{dbltopnumber}{5}
\renewcommand{\topfraction}{0.95}
\renewcommand{\dbltopfraction}{0.95}
\renewcommand{\textfraction}{0.05}
\renewcommand{\floatpagefraction}{0.9}
\renewcommand{\dblfloatpagefraction}{0.9}
\newcommand{\IncludePGF}[1]{\resizebox{\linewidth}{!}{\input{#1}}}

\ifPDFTeX
\DeclareUnicodeCharacter{2013}{\textendash}
\DeclareUnicodeCharacter{2014}{\textemdash}
\DeclareUnicodeCharacter{2018}{\textquoteleft}
\DeclareUnicodeCharacter{2019}{\textquoteright}
\DeclareUnicodeCharacter{201C}{\textquotedblleft}
\DeclareUnicodeCharacter{201D}{\textquotedblright}
\DeclareUnicodeCharacter{2212}{\textminus}
\fi

\pdfstringdefDisableCommands{%
    \def\vct#1{#1}%
    \def\mat#1{#1}%
    \def\matg#1{#1}%
    \def\I{I}%
}

\newtheoremstyle{bodysmall}{3pt}{3pt}{\normalfont\small}{0pt}{\bfseries\small}{.}{0.5em}{}
\theoremstyle{bodysmall}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem*{theorem*}{Theorem}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\newcommand{\Skew}[1]{\left[#1\right]_\times}
\newcommand{\Id}{\mat{I}}
\newcommand{\pct}[1]{#1\%}
\newcommand{\meanstd}[2]{#1 \pm #2}
\newcommand{\ci}[2]{#1\,(\,#2\,\text{CI}\,)}
\newcommand{\Ito}{\mathrm{It\hat{o}}}
\newcommand{\ItoInt}[2]{\int_0^{#1} #2\, dW_s^{\Ito}}
\DeclareMathOperator{\proj}{\mathcal{P}}
\DeclareMathOperator{\clip}{clip}
\DeclareMathOperator{\wrap}{wrap}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\blkdiag}{blkdiag}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\sym}{sym}
\newcommand{\set}[1]{\mathcal{#1}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\Cov}{Cov}
\newcommand{\ts}[1]{\text{#1}}
\newcommand{\fitcol}[1]{\adjustbox{max width=\columnwidth}{\ensuremath{\displaystyle #1}}}

\makeatletter
\@ifpackageloaded{cite}{\renewcommand{\citepunct}{,\penalty\citepunctpenalty\,}\renewcommand{\citedash}{--}}{}
\makeatother
\emergencystretch=2em
\allowdisplaybreaks

\title{OU-Driven Quaternion MEKF for Marine INS and Wave-State Estimation (draft)}
\author{%
    \IEEEauthorblockN{Mikhail S. Grushinskiy}
    \IEEEauthorblockA{Independent Researcher, USA, Fair Lawn\\
    Bareboat Necessities GitHub Project\\
    Email: mgrouch@users.github.com}%
}
\pgfplotsset{compat=1.18}

\begin{document}
\maketitle

\begin{abstract}
We present a quaternion multiplicative EKF for ocean-wave motion that couples
attitude to a linear kinematic chain for velocity, oscillatory displacement,
and the integral of displacement. World-frame acceleration is assigned an
Ornstein--Uhlenbeck prior with stationary variance and finite temporal
correlation. Gyroscope and accelerometer biases are modeled explicitly. To
suppress integration drift, an anisotropic zero pseudo-measurement is applied
to the integral state. We derive the continuous model, analytic OU-chain
discretization, numerically stable small-step covariance formulas, and explicit
accelerometer and magnetometer Jacobians. A lightweight estimator obtains the
horizontal wave-propagation axis and apparent travel sense from leveled
acceleration. Online adaptation tunes the OU time scale, stationary acceleration
scale, and the standard-deviation scale whose square forms the integral
pseudo-measurement covariance. The cubic scaling with correlation time is a
physically motivated engineering normalization for fixed update cadence and a
bounded family of spectra. The
method is suitable for real-time deployment with low-cost embedded IMUs.
\end{abstract}

\begin{IEEEkeywords}
extended Kalman filter, inertial navigation, IMU, marine Motion Reference Unit, MRU, Ornstein--Uhlenbeck process, Heave Reference System, HRS, heave sensor, wave height estimation, wave measurement buoy\end{IEEEkeywords}

\input{w3d-intro.tex-part}
\input{w3d-state.tex-part}
\input{w3d-meas.tex-part}
\input{w3d-wind-heel.tex-part}
\input{w3d-proc-cont.tex-part}
\input{w3d-lti-discrete.tex-part}
\input{w3d-analytic-coeff.tex-part}
\input{w3d-kalm-up.tex-part}
\input{w3d-init.tex-part}
\input{w3d-fus-methods.tex-part}
\input{w3d-sim-charts.tex-part}
\input{w3d-conclusion-summary.tex-part}
\input{w3d-gps-fusion.tex-part}

\FloatBarrier
\section*{Acknowledgments}
This work was prepared with the supervised assistance of artificial intelligence tools, including ChatGPT, Claude and DeepSeek, which were employed to aid in code generation, mathematical drafting, and editorial preparation. All conceptual development, validation of results, and final revisions were conducted independently by the author.

\section*{Disclosure Statement}
The author declares no conflicts of interest and received no funding for this research.

\nocite{Kalman1960_Filtering,Jazwinski1970_Filtering,Maybeck1979_StochasticModels,Markley2003AttitudeError,UhlenbeckOrnstein1930_OU}
\bibliographystyle{IEEEtran}
\bibliography{w3d}
\balance
\end{document}
