% ============================================================
%  Applying Vessel RAOs to 3D PM/JONSWAP Wave Simulations
%  IEEE Two-Column Journal Format
% ============================================================
\documentclass[journal,10pt,twocolumn]{IEEEtran}

\usepackage{iftex}
\ifPDFTeX
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\fi

\usepackage{amsmath,amsfonts,amssymb,bm,mathtools}
\usepackage{newtxtext,newtxmath}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage[breaklinks=true,hidelinks]{hyperref}
\usepackage[nameinlink,capitalise,noabbrev]{cleveref}
\usepackage[final]{microtype}

\interdisplaylinepenalty=2500
\allowdisplaybreaks

% --- Math shorthands ---
\newcommand{\dd}{\,\mathrm{d}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\ee}{\mathrm{e}}
\newcommand{\Real}{\operatorname{Re}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\vect}[1]{\bm{#1}}

% --- Custom column types ---
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% ============================================================
\begin{document}

    \title{Applying Vessel Response Amplitude Operators\\
    to Existing Three-Dimensional Wave Simulations}

    \author{%
        \IEEEauthorblockN{Mikhail Grushinskiy}\\
        \IEEEauthorblockA{Bareboat Necessities Project}
    }

    \maketitle

% ============================================================
    \begin{abstract}
        A three-dimensional irregular wave simulation generated from the Pierson--Moskowitz
        (PM) or JONSWAP spectrum describes the incident sea state but does not by itself yield
        vessel behavior.
        The missing bridge is the vessel hydrodynamic transfer model, which in linear seakeeping
        is most compactly expressed as a set of complex-valued response amplitude operators
        (RAOs), or more fundamentally as frequency-dependent added mass, radiation damping,
        hydrostatic restoring, and wave-excitation coefficients.
        This paper gives a self-contained workflow for coupling an existing 3D directional
        wave realization to a linear vessel seakeeping model.
        Topics covered include directional wave-field decomposition, component-by-component
        application of complex RAOs to recover phase-consistent motion time series, response
        spectra and root-mean-square (RMS) statistics, encounter frequency and forward-speed
        effects, kinematic stretching for sub-surface loads, sea-state disturbance around the
        hull, extension to the time-domain Cummins equation with radiation memory, state-space
        realization of the convolution kernel, and a structured nine-step practical algorithm
        with emphasis on phase, heading, amplitude, and frequency-coverage pitfalls.
        The scope and limitations of linear RAO methods relative to time-domain potential-flow
        and CFD approaches are assessed, and a comparison of common RAO source methods is
        provided.
    \end{abstract}

    \begin{IEEEkeywords}
        Seakeeping, response amplitude operator, RAO, PM spectrum, JONSWAP,
        irregular waves, vessel motions, wave excitation, Cummins equation,
        radiation memory, state-space, directional spreading, Wheeler stretching,
        offshore engineering.
    \end{IEEEkeywords}

% ============================================================
    \section{Introduction}
    \label{sec:intro}

    Suppose an irregular free surface has already been synthesized in three dimensions from a
    spectral model such as Pierson--Moskowitz (PM) or JONSWAP, producing a realization
    $\eta(\vect{r},t)$ that is spatially and temporally coherent, reproduces directional
    spreading, random phase, and realistic wave superposition.

    A vessel does not respond to the free surface alone.
    Vessel behavior depends on the interaction between the incoming wave field and the hull
    geometry.
    In linear seakeeping this interaction is captured by \emph{response amplitude operators}
    (RAOs)~\cite{Faltinsen,Newman}: complex transfer functions that map a unit-amplitude
    regular incident wave at a given frequency and heading to a steady-state response such
    as heave, pitch, roll, deck acceleration, relative motion, or structural load.

    The essential computational pipeline is:
    \begin{equation}
        \boxed{
            \substack{\displaystyle\text{incident}\\[1pt]\displaystyle\text{wave field}}
            \;\xrightarrow{\;\text{decompose}\;}\;
            \substack{\displaystyle\text{directional}\\[1pt]\displaystyle\text{components}}
            \;\xrightarrow{\;\text{RAO}\;}\;
            \substack{\displaystyle\text{motions /}\\[1pt]\displaystyle\text{loads}}
        }
        \label{eq:pipeline}
    \end{equation}

    If the vessel transfer model is known, the existing PM/JONSWAP realization can be
    converted into vessel response via two main routes:
    \begin{enumerate}[label=(\roman*),leftmargin=1.4em]
    \item \textbf{Frequency-domain (spectral) route:} multiply the directional wave
    spectrum by $|H|^2$ to obtain response spectra and RMS statistics directly.
    \item \textbf{Time-domain route:} apply the complex RAO to each wave component and
    sum the resulting contributions, preserving phase coherence with the original
    realization.
    \end{enumerate}
    Both routes are explained in detail in the sections that follow.
    When mooring coupling, control feedback, or transient dynamics are important, a
    deeper time-domain model based on the Cummins equation is the appropriate extension and
    is treated in \cref{sec:cummins}.

% ============================================================
    \section{What the Wave Simulation Already Contains}
    \label{sec:wavemodel}

    \subsection{Directional Spectrum}
    \label{sec:dirspectrum}

    Most irregular-wave simulations begin from a directional variance spectrum
    \begin{equation}
        S_\eta(\omega,\mu),
    \end{equation}
    where $\omega$ is angular frequency and $\mu$ is wave propagation direction.
    The directional spectrum is commonly factorized as
    \begin{equation}
        S_\eta(\omega,\mu) = S_0(\omega)\,D(\omega,\mu),
        \label{eq:factorized}
    \end{equation}
    where $S_0(\omega)$ is a one-dimensional frequency spectrum and $D(\omega,\mu)$ is a
    directional spreading function satisfying
    \begin{equation}
        \int_{-\pi}^{\pi} D(\omega,\mu)\dd\mu = 1.
    \end{equation}

    For the PM spectrum, the one-dimensional form is
    \begin{equation}
        S_0(\omega) = \frac{5}{16}H_s^2\omega_p^4\omega^{-5}
        \exp\!\left[-\frac{5}{4}\!\left(\frac{\omega_p}{\omega}\right)^{\!4}\right],
        \label{eq:PM}
    \end{equation}
    and the JONSWAP spectrum augments \eqref{eq:PM} by a peak enhancement factor~$\gamma$:
    \begin{equation}
        S_{\mathrm{JS}}(\omega) = S_0(\omega)\cdot
        \gamma^{\exp\!\left[-\frac{(\omega-\omega_p)^2}{2\sigma^2\omega_p^2}\right]},
        \label{eq:JONSWAP}
    \end{equation}
    with $\gamma\approx3.3$ and $\sigma=0.07$ ($\omega\le\omega_p$),
    $\sigma=0.09$ ($\omega>\omega_p$).

    A common directional spreading function is the cosine-power model
    \begin{equation}
        D(\mu) = C_s\cos^{2s}\!\left(\frac{\mu-\mu_0}{2}\right),
        \label{eq:spreading}
    \end{equation}
    where $\mu_0$ is the mean direction, $s$ controls angular spread, and $C_s$ is a
    normalizing constant.
    Higher values of $s$ produce narrower, more uni-directional seas.

    \subsection{Free-Surface Realization}
    \label{sec:realization}

    The 3D wave field is synthesized as a superposition of directional sinusoidal
    components:
    \begin{equation}
        \eta(\vect{r},t)
        = \sum_{m=1}^{M}
        a_m\cos\!\bigl(\vect{k}_m\!\cdot\!\vect{r} - \omega_m t + \phi_m\bigr),
        \label{eq:realization}
    \end{equation}
    where $\vect{r}=[x\;\;y]^T$, $\vect{k}_m$ is the horizontal wavevector of component
    $m$, $\phi_m\sim\mathcal{U}(0,2\pi)$ is a random phase, and the component amplitude is
    \begin{equation}
        a_m = \sqrt{2\,S_\eta(\omega_m,\mu_m)\,\Delta\omega\,\Delta\mu}.
        \label{eq:amplitude}
    \end{equation}

    \subsection{Dispersion and Wavevector}
    \label{sec:dispersion}

    Wavenumber $k_m=|\vect{k}_m|$ is obtained from the linear dispersion relation.
    For finite depth $h$,
    \begin{equation}
        \omega_m^2 = g k_m\tanh(k_m h),
        \label{eq:dispersion}
    \end{equation}
    which reduces to $k_m=\omega_m^2/g$ in deep water.
    The wavevector corresponding to propagation direction $\mu_m$ is
    \begin{equation}
        \vect{k}_m = k_m\begin{bmatrix}\cos\mu_m\\\sin\mu_m\end{bmatrix}.
        \label{eq:wavevec}
    \end{equation}

    \subsection{What Is Already Available}
    \label{sec:available}

    A pre-existing PM/JONSWAP 3D synthesis already provides everything needed to drive a
    linear vessel model:
    \begin{enumerate}[label=(\arabic*),leftmargin=1.4em]
    \item frequencies $\{\omega_m\}$,
    \item directions $\{\mu_m\}$,
    \item wavevectors $\{\vect{k}_m\}$,
    \item component amplitudes $\{a_m\}$,
    \item random phases $\{\phi_m\}$,
    \item spatial phase via $\vect{k}_m\cdot\vect{r}$.
    \end{enumerate}
    The free-surface field alone does not, however, determine vessel loads.
    Hull geometry and fluid--structure interaction must be supplied separately through a
    hydrodynamic model.

% ============================================================
    \section{Response Amplitude Operators}
    \label{sec:rao}

    \subsection{Definition and Properties}
    \label{sec:raodef}

    For any linear response quantity $q$, the RAO is a complex transfer function
    \begin{equation}
        H_q(\omega,\beta) = |H_q(\omega,\beta)|\,\ee^{\ii\varphi_q(\omega,\beta)},
        \label{eq:raodef}
    \end{equation}
    where $\omega$ is wave (or encounter) frequency, $\beta$ is relative wave heading with
    respect to the vessel, and $\varphi_q$ is the phase lead of the response with respect
    to the incident wave crest at the reference point.

    If a unit-amplitude regular wave excites the vessel, the steady-state response is
    \begin{equation}
        q(t) = \Real\!\left\{H_q(\omega,\beta)\,\ee^{\ii\omega t}\right\}
        = |H_q|\cos(\omega t + \varphi_q).
    \end{equation}
    Thus the RAO captures amplitude gain and phase shift simultaneously.

    \subsection{The Six Rigid-Body Motions}
    \label{sec:6dof}

    Let the vessel-fixed body frame have $x_b$ forward, $y_b$ to port, and $z_b$ upward.
    The six rigid-body motions are
    \begin{equation}
    \begin{aligned}
        \vect{\xi}
        &= \begin{bmatrix}
              \xi_1 & \xi_2 & \xi_3 & \xi_4 & \xi_5 & \xi_6
        \end{bmatrix}^T\\
        &= \begin{bmatrix}
              \text{surge} & \text{sway} & \text{heave} &
              \text{roll}  & \text{pitch} & \text{yaw}
        \end{bmatrix}^T.
        \label{eq:dofs}
    \end{aligned}
    \end{equation}
    A separate RAO $H_j(\omega,\beta)$ is defined for each DOF~$j=1,\ldots,6$.

    \subsection{Relative Heading}
    \label{sec:heading}

    If the absolute wave propagation direction is $\mu$ and the vessel heading is $\psi$,
    then the relative wave heading used to look up the RAO is
    \begin{equation}
        \beta = \mu - \psi,
        \label{eq:beta}
    \end{equation}
    modulo any sign convention in the hydrodynamic software.
    Common conventions include: $\beta=180^\circ$ for head seas (waves from ahead),
    $\beta=0^\circ$ for following seas, and $\beta=90^\circ$ for beam seas from
    starboard.
    This must be verified against the specific RAO data source.

    \subsection{Units and Normalization}
    \label{sec:units}

    \Cref{tab:raounits} summarizes typical RAO units by output type.
    The single most common source of error is a factor-of-two mismatch between
    amplitude-normalized ($a=H/2$) and height-normalized ($H$) RAO conventions.

    \begin{table}[t]
        \centering
        \caption{Typical RAO Units by Response Quantity}
        \label{tab:raounits}
        \small
        \begin{tabular}{@{}lc@{}}
            \toprule
            Response quantity & RAO units \\
            \midrule
            Surge, sway, heave       & m/m \\
            Roll, pitch, yaw         & rad/m \;\text{or}\; deg/m \\
            Translational velocity   & (m/s)/m \\
            Translational acceleration & (m/s$^2$)/m \\
            Angular acceleration     & (rad/s$^2$)/m \\
            Structural moment        & (N$\cdot$m)/m \\
            Relative motion          & m/m \\
            \bottomrule
        \end{tabular}
    \end{table}

    \subsection{Sources of RAO Data}
    \label{sec:raosources}

    RAOs may be obtained from several methodologies, summarized in \cref{tab:raosources}.
    Potential-flow panel codes such as WAMIT~\cite{WAMIT}, AQWA, OrcaWave, NEMOH, and
    Capytaine~\cite{Capytaine} solve the boundary-value problem for the radiation and
    diffraction potentials and output frequency-dependent added mass, radiation damping,
    excitation forces, and motion RAOs.
    Strip theory codes provide faster but less accurate solutions for slender
    ship-like hulls.
    Physical model tests remain the gold standard for complex geometries and validation.

    \begin{table}[t]
        \centering
        \caption{RAO Source Methods: Scope and Considerations}
        \label{tab:raosources}
        \small
        \begin{tabular}{@{}L{1.8cm}L{1.6cm}L{2.4cm}@{}}
            \toprule
            Method & Typical scope & Key limitations \\
            \midrule
            3D panel (WAMIT, AQWA, Capytaine)
            & All hull types, 6-DOF
            & Potential flow only; viscous roll damping requires correction \\
            Strip theory
            & Slender ships, forward speed
            & Poor for short waves, offshore units, or large beam/draft \\
            Model tests
            & Any hull, nonlinear effects
            & Cost, scale effects, limited sea-state coverage \\
            CFD / RANS
            & Extreme events, slamming
            & Very high computational cost; rarely used for RAO databases \\
            \bottomrule
        \end{tabular}
    \end{table}

% ============================================================
    \section{Applying RAOs to Wave Components}
    \label{sec:applyrao}

    \subsection{Stationary or Slowly Moving Vessel}
    \label{sec:stationary}

    Assume the vessel is stationary in mean position, or moving slowly enough that
    forward-speed effects are negligible.
    Let $\vect{r}_0$ be the vessel reference point (e.g., center of gravity or RAO
    reference origin).

    Each incident wave component $m$ from \eqref{eq:realization} generates a response
    contribution
    \begin{equation}
        q_m(t)
        = \Real\!\left\{
        H_q(\omega_m,\beta_m)\,a_m\,
        \ee^{\ii(\vect{k}_m\cdot\vect{r}_0 - \omega_m t + \phi_m)}
        \right\}.
        \label{eq:component_response}
    \end{equation}
    The total response is the superposition
    \begin{equation}
        q(t)
        = \sum_{m=1}^{M} q_m(t)
        = \Real\!\left\{
        \sum_{m=1}^{M}
        H_q(\omega_m,\beta_m)\,a_m\,
        \ee^{\ii(\vect{k}_m\cdot\vect{r}_0 - \omega_m t + \phi_m)}
        \right\}.
        \label{eq:response_sum}
    \end{equation}
    \Cref{eq:response_sum} is the central formula of the paper.
    The RAO acts on each component independently, scaling its amplitude by $|H_q|$ and
    shifting its phase by $\varphi_q$, and preserves phase coherence with the original wave
    realization.

    For all six rigid-body DOFs simultaneously,
    \begin{equation}
        \xi_j(t)
        = \Real\!\left\{
        \sum_{m=1}^{M}
        H_j(\omega_m,\beta_m)\,a_m\,
        \ee^{\ii(\vect{k}_m\cdot\vect{r}_0 - \omega_m t + \phi_m)}
        \right\},
        \label{eq:motion_dof}
    \end{equation}
    computed once for each $j=1,\ldots,6$ from the same component list.

    \subsection{Encounter Frequency and Forward Speed}
    \label{sec:encounter}

    If the vessel moves with mean horizontal velocity $\vect{U}$, the phase of component
    $m$ at the vessel is
    \begin{equation}
        \theta_m(t) = \vect{k}_m\cdot\vect{r}_0(t) - \omega_m t + \phi_m,
    \end{equation}
    and the rate of phase change is the encounter frequency
    \begin{equation}
        \omega_{e,m} = \omega_m - \vect{k}_m\cdot\vect{U}.
        \label{eq:encounter}
    \end{equation}
    For scalar speed $U$ and angle $\chi_m$ between wave propagation and vessel velocity,
    \begin{equation}
        \omega_{e,m} = \omega_m - k_m U\cos\chi_m.
        \label{eq:encounter_scalar}
    \end{equation}

    When forward speed is significant, the RAO must be evaluated at encounter frequency:
    \begin{equation}
        q(t)
        = \Real\!\left\{
        \sum_{m=1}^{M}
        H_q(\omega_{e,m},\beta_m)\,a_m\,
        \ee^{\ii(\vect{k}_m\cdot\vect{r}_0(t) - \omega_m t + \phi_m)}
        \right\}.
        \label{eq:response_encounter}
    \end{equation}
    For a moored or stationary vessel, $\vect{U}=\vect{0}$ and $\omega_e=\omega$.

% ============================================================
    \section{Spectral Route: Response Spectra and RMS Statistics}
    \label{sec:spectral}

    \subsection{Single-Output Response Spectrum}
    \label{sec:responsespec}

    If only second-order statistics (RMS, variance, response spectrum) are required, it is
    more efficient to work in the frequency domain.
    For RAO $H_q(\omega,\beta)$ and directional wave spectrum $S_\eta(\omega,\mu)$, the
    one-sided response spectrum is
    \begin{equation}
        S_q(\omega)
        = \int_{-\pi}^{\pi}
        |H_q(\omega_e,\beta)|^2\,S_\eta(\omega,\mu)\dd\mu,
        \label{eq:response_spectrum}
    \end{equation}
    where $\beta$ and $\omega_e$ follow from $\mu$, $\psi$, and $\vect{U}$ as in
    \cref{sec:encounter}.
    The response variance and RMS are
    \begin{equation}
        \sigma_q^2 = \int_0^\infty S_q(\omega)\dd\omega,
        \qquad
        q_{\mathrm{RMS}} = \sigma_q.
        \label{eq:rms}
    \end{equation}

    \subsection{Cross-Spectra and Covariance}
    \label{sec:crossspec}

    For two outputs $q_1$ and $q_2$, the cross-spectrum is
    \begin{equation}
        S_{q_1q_2}(\omega)
        = \int_{-\pi}^{\pi}
        H_{q_1}(\omega_e,\beta)\,
        H_{q_2}^{*}(\omega_e,\beta)\,
        S_\eta(\omega,\mu)\dd\mu.
        \label{eq:crossspectrum}
    \end{equation}
    The real part is the co-spectrum and the imaginary part is the quadrature spectrum.
    The full covariance matrix of vessel motions is assembled from the six-by-six
    cross-spectral matrix, which is useful for multi-axis operability studies and combined
    load effects.

    \subsection{Spectral Moments and Derived Statistics}
    \label{sec:moments}

    Spectral moments of order $n$ are
    \begin{equation}
        m_n = \int_0^\infty \omega^n S_q(\omega)\dd\omega.
    \end{equation}
    Key derived statistics include:
    \begin{itemize}[leftmargin=1.2em]
        \item significant amplitude: $q_{1/3} = 4\sqrt{m_0}$,
        \item mean upcrossing period: $T_{02} = 2\pi\sqrt{m_0/m_2}$,
        \item spectral bandwidth: $\varepsilon = \sqrt{1 - m_2^2/(m_0 m_4)}$,
        \item most probable maximum in $N$ cycles: $q_{\max}\approx\sigma_q\sqrt{2\ln N}$
        (narrow-band approximation).
    \end{itemize}

% ============================================================
    \section{Wave Kinematics and Wheeler Stretching}
    \label{sec:stretching}

    \subsection{Sub-Surface Kinematics From Linear Theory}
    \label{sec:kinematics}

    For load calculations on slender members via the Morison equation~\cite{Morison}, the
    horizontal fluid velocity and acceleration at position $(x,y,z)$ are required in
    addition to the surface elevation.
    Under Airy theory, for component $m$ at depth $z$ ($-h\le z\le0$),
    \begin{align}
        u_m &= a_m\omega_m
        \frac{\cosh[k_m(z+h)]}{\sinh(k_m h)}
        \cos(\vect{k}_m\cdot\vect{r}-\omega_m t+\phi_m), \\
        \dot{u}_m &= a_m\omega_m^2
        \frac{\cosh[k_m(z+h)]}{\sinh(k_m h)}
        \sin(\vect{k}_m\cdot\vect{r}-\omega_m t+\phi_m).
        \label{eq:kinematics}
    \end{align}
    Linear theory only provides kinematics up to the mean water level $z=0$; kinematics
    between the mean level and the actual free surface $\eta$ must be obtained by
    extrapolation or stretching.

    \subsection{Wheeler Stretching}
    \label{sec:wheeler}

    Wheeler stretching~\cite{Wheeler} is the standard engineering method for extending
    Airy kinematics into the wave crest region.
    The stretching replaces the physical depth coordinate $z$ with a mapped coordinate
    $z'$ defined such that the argument of the hyperbolic functions is evaluated at
    \begin{equation}
        z' = \frac{h(z-\eta)}{h+\eta},
        \label{eq:wheeler}
    \end{equation}
    so that the linear kinematics formula is re-evaluated at $z'$ instead of $z$.
    This maps the interval $[-h,\eta]$ bijectively onto $[-h,0]$, ensuring that the
    kinematics remain finite at the instantaneous free surface without the exponential
    blow-up that occurs if the linear formula is applied directly above $z=0$.

    Applying Wheeler stretching to \eqref{eq:kinematics}, the horizontal velocity at
    physical height $z$ (for $-h\le z\le\eta$) becomes
    \begin{equation}
        u_m^{\mathrm{W}}
        = a_m\omega_m
        \frac{\cosh[k_m(z'+h)]}{\sinh(k_m h)}
        \cos(\vect{k}_m\cdot\vect{r}-\omega_m t+\phi_m),
        \label{eq:wheeler_u}
    \end{equation}
    with $z'$ from \eqref{eq:wheeler}.
    Wheeler stretching is widely used in Morison load calculations on jacket structures and
    monopiles under steep irregular waves and is specified in several industry codes
    (DNV-RP-C205~\cite{DNVGL}).

% ============================================================
    \section{Outputs Beyond the Six Rigid-Body Motions}
    \label{sec:outputs}

    \subsection{Arbitrary Point Kinematics}
    \label{sec:pointmotion}

    For a point $P$ fixed in the body at offset $\vect{r}_{GP}$ from the center of gravity
    $G$, the small-angle translational displacement is
    \begin{equation}
        \vect{u}_P(t) \approx \vect{u}_G(t) + \vect{\theta}(t)\times\vect{r}_{GP},
        \label{eq:point_disp}
    \end{equation}
    and the acceleration is
    \begin{equation}
        \vect{a}_P(t) \approx \ddot{\vect{u}}_G(t) + \ddot{\vect{\theta}}(t)\times\vect{r}_{GP}.
        \label{eq:point_accel}
    \end{equation}
    Once the six-DOF motion time series are available from \eqref{eq:motion_dof}, these
    expressions yield accelerations at any sensor location, crane tip, bridge deck, or
    moonpool.
    Many hydrodynamic tools also provide point-acceleration RAOs directly.

    \subsection{Relative Motion and Air Gap}
    \label{sec:relmotion}

    Relative vertical motion between a point $P$ on the vessel and the local wave elevation
    is a critical operability metric for hatches, moonpools, and helicopter decks.
    The relative motion RAO is
    \begin{equation}
        H_{\mathrm{rel}}(\omega,\beta)
        = H_3(\omega,\beta)
        - x_P H_5(\omega,\beta)
        + y_P H_4(\omega,\beta)
        - K_p(z_P),
        \label{eq:relmotion}
    \end{equation}
    where $H_3$, $H_4$, $H_5$ are heave, roll, and pitch RAOs, $(x_P,y_P)$ is the
    horizontal offset of point $P$, and $K_p$ is the wave pressure response factor at the
    undisturbed surface.
    Significant relative motion $s_{1/3}=4\sqrt{m_{0,\mathrm{rel}}}$ is used to evaluate
    freeboard adequacy.

    \subsection{Sea-State Disturbance Near the Hull}
    \label{sec:disturbance}

    The incident wave field is modified in the near-field of the hull by scattering and
    radiation.
    For objects close to the hull (risers, gangways, moonpool equipment, adjacent barges),
    the locally disturbed wave elevation and velocity differ from the undisturbed incident
    field.
    Tools such as OrcaWave and WAMIT output \emph{sea-state disturbance RAOs}: complex,
    position-, frequency-, and direction-dependent operators for the disturbed velocity
    potential or surface elevation, applicable within a few vessel beam lengths of the hull.
    These allow the same PM/JONSWAP component set to be used to recover the disturbed fluid
    kinematics at any near-vessel location.

% ============================================================
    \section{From Wave Grid to Response: Practical Reconstruction}
    \label{sec:reconstruction}

    \subsection{Best Case: Original Component List Retained}
    \label{sec:bestcase}

    The preferred workflow retains the explicit list
    $\{\omega_m,\mu_m,\vect{k}_m,a_m,\phi_m\}_{m=1}^{M}$
    from the wave synthesis stage.
    Each component is then passed through the RAO as in
    \cref{eq:component_response,eq:response_sum}, giving a response that is
    \emph{phase-consistent} with the incident sea realization: if the synthesis used a
    specific random seed, the vessel motion and the surface elevation share a unique
    stochastic realization rather than merely the same spectral density.

    \subsection{If Only the Wave Grid Is Available}
    \label{sec:gridrec}

    If the original component list is unavailable but the free surface $\eta(x,y,t)$ is
    known on a grid, directional spectral content can be recovered by a spatio-temporal
    Fourier transform
    \begin{equation}
        \hat{\eta}(k_x,k_y,\omega)
        = \int\!\!\int\!\!\int
        \eta(x,y,t)\,
        \ee^{-\ii(k_x x+k_y y-\omega t)}\dd x\,\dd y\,\dd t,
    \end{equation}
    from which approximate directional spectrum estimates and surrogate wave components can
    be synthesized.
    The resulting response will be statistically consistent with the spectrum but will not
    share the exact phase realizations of the original simulation.
    Whenever possible, the component list should be retained from the wave generation stage.

% ============================================================
    \section{The Time-Domain Route: Cummins Equation}
    \label{sec:cummins}

    \subsection{Motivation}
    \label{sec:cummins_why}

    Direct RAO superposition is ideal for linear response prediction.
    A time-domain model is needed when mooring-line or riser coupling is strong, control
    systems or station-keeping are modeled, transient or start-up dynamics matter, multiple
    coupled subsystems (crane, payload, adjacent vessel) are present, or nonlinear damping
    and mooring restoring must be included.
    The same PM/JONSWAP incident wave components drive both approaches.

    \subsection{Cummins Equation}
    \label{sec:cummins_eq}

    The standard linear potential-flow time-domain model is the Cummins equation~\cite{Cummins}:
    \begin{multline}
        \bigl(\vect{M}+\vect{A}_\infty\bigr)\ddot{\vect{\xi}}(t)
        + \int_0^t \vect{K}_r(t-\tau)\dot{\vect{\xi}}(\tau)\dd\tau \\
        + \vect{C}_h\vect{\xi}(t)
        = \vect{F}_{\mathrm{exc}}(t) + \vect{F}_{\mathrm{ext}}(t),
        \label{eq:cummins}
    \end{multline}
    where $\vect{M}$ is the rigid-body mass/inertia matrix, $\vect{A}_\infty$ is the
    added mass at infinite frequency, $\vect{K}_r(t)$ is the radiation impulse-response
    (retardation) kernel, $\vect{C}_h$ is the hydrostatic restoring matrix,
    $\vect{F}_{\mathrm{ext}}$ collects mooring, damping, control, and thruster forces,
    and $\vect{F}_{\mathrm{exc}}(t)$ is the first-order wave excitation force vector.

    The retardation kernel is related to the frequency-domain radiation damping matrix
    $\vect{B}(\omega)$ by
    \begin{equation}
        \vect{K}_r(t)
        = \frac{2}{\pi}\int_0^\infty \vect{B}(\omega)\cos(\omega t)\dd\omega,
        \label{eq:retardation}
    \end{equation}
    and equivalently $\vect{A}(\omega) = \vect{A}_\infty
    - (1/\omega)\int_0^\infty \vect{K}_r(t)\sin(\omega t)\dd t$.

    \subsection{Wave Excitation Force From the Component List}
    \label{sec:excitation}

    The wave excitation force vector is built from the same PM/JONSWAP incident wave
    components:
    \begin{equation}
        \vect{F}_{\mathrm{exc}}(t)
        = \Real\!\left\{
        \sum_{m=1}^{M}
        \vect{X}(\omega_{e,m},\beta_m)\,a_m\,
        \ee^{\ii(\vect{k}_m\cdot\vect{r}_0(t) - \omega_m t + \phi_m)}
        \right\},
        \label{eq:excitation}
    \end{equation}
    where $\vect{X}(\omega,\beta)$ is the complex vector of wave-excitation force and moment
    coefficients obtained from the panel code.
    Vessel motion is then obtained by integrating \eqref{eq:cummins} rather than by
    applying a direct motion RAO.

    \subsection{State-Space Realization of Radiation Memory}
    \label{sec:statespace}

    Evaluating the convolution integral in \eqref{eq:cummins} at every time step is
    computationally costly for large~$M$.
    A standard remedy is to fit a minimal-order linear state-space model to
    $\vect{K}_r(t)$~\cite{Perez}:
    \begin{align}
        \dot{\vect{z}}(t) &= \vect{A}_r\vect{z}(t) + \vect{B}_r\dot{\vect{\xi}}(t),
        \label{eq:ss_state} \\
        \vect{\mu}(t)     &= \vect{C}_r\vect{z}(t),
        \label{eq:ss_output}
    \end{align}
    where $\vect{z}$ is the internal state vector and $\vect{\mu}(t)$ approximates the
    radiation force convolution.
    Replacing the integral in \eqref{eq:cummins} with \eqref{eq:ss_output} yields an
    ordinary differential equation that is numerically efficient to integrate.
    The system matrices $(\vect{A}_r,\vect{B}_r,\vect{C}_r)$ are fitted by balanced
    realization or frequency-domain least-squares using the Bode response of $\vect{K}_r$.
    Packages such as MSS Toolbox~\cite{MSS} and HydroDyn~\cite{HydroDyn} provide
    implementations of this fitting procedure.

% ============================================================
    \section{A Practical Nine-Step Algorithm}
    \label{sec:algorithm}

    Assume a 3D directional wave field has been synthesized from PM or JONSWAP.

    \paragraph{Step 1: Retain the wave component list.}
    Preserve $\{\omega_m,\mu_m,\vect{k}_m,a_m,\phi_m\}_{m=1}^M$ from the synthesis stage.

    \paragraph{Step 2: Obtain vessel RAOs.}
    For each output of interest, obtain a complex RAO table $H_q(\omega,\beta)$ from a
    panel code, strip-theory code, or model test.
    Confirm the amplitude normalization (per wave amplitude or per wave height), heading
    sign convention, phase reference point, and sign convention for rotations.

    \paragraph{Step 3: Define vessel mean pose and speed.}
    Specify heading $\psi$ and, if applicable, mean velocity $\vect{U}$.

    \paragraph{Step 4: Verify RAO frequency and heading coverage.}
    The RAO table must cover the full frequency range of $S_\eta(\omega,\mu)$ with
    sufficient resolution.
    Gaps at low or high frequencies cause underprediction of response.
    Coverage in heading must span at least the dominant directions present in the sea state.

    \paragraph{Step 5: Compute relative headings.}
    For each component, compute $\beta_m=\mu_m-\psi$ modulo the software heading convention.

    \paragraph{Step 6: Compute encounter frequency.}
    If forward speed matters, compute $\omega_{e,m}=\omega_m-\vect{k}_m\cdot\vect{U}$.

    \paragraph{Step 7: Interpolate the complex RAO.}
    Evaluate $H_q(\omega_{e,m},\beta_m)$ by bilinear or spline interpolation on the
    $(\omega,\beta)$ grid.
    Interpolate real and imaginary parts separately rather than amplitude and wrapped phase.

    \paragraph{Step 8: Synthesize the response.}
    Compute the time series
    \begin{equation}
        q(t) = \Real\!\left\{
        \sum_{m=1}^{M}
        H_q(\omega_{e,m},\beta_m)\,a_m\,
        \ee^{\ii(\vect{k}_m\cdot\vect{r}_0(t)-\omega_m t+\phi_m)}
        \right\}.
        \label{eq:final_synthesis}
    \end{equation}

    \paragraph{Step 9: Post-process.}
    Derive RMS values, spectral moments, significant amplitudes, peak estimates, crossing
    rates, operability metrics, or fatigue damage equivalent loads as required.

% ============================================================
    \section{Implementation Pitfalls}
    \label{sec:pitfalls}

    \subsection{Amplitude Convention Mismatch}
    \label{sec:amp_mismatch}

    If the wave simulation uses component amplitude $a=H/2$ but the RAO is tabulated per
    wave height $H$, responses will be overestimated by a factor of two.
    Verify the normalization before any computation.

    \subsection{Phase Convention Mismatch}
    \label{sec:phase_mismatch}

    Phase conventions are not standardized across tools.
    Key questions to resolve: Is the RAO phase a lead or a lag? Is the time convention
    $\ee^{+\ii\omega t}$ or $\ee^{-\ii\omega t}$? What is the wave reference point for
    phase (e.g., the coordinate origin, center of gravity, or waterplane centroid)? What
    sign convention is used for rotations?
    A mismatch here changes the response phase but not the amplitude spectrum.
    Validation against a known regular-wave case should always be performed.

    \subsection{Heading Convention Mismatch}
    \label{sec:head_mismatch}

    $\beta=0^\circ$ is not universally defined; different codes use head seas, following
    seas, or wave from north.
    Applying the RAO with a heading offset of $180^\circ$ gives the wrong response for
    asymmetric motions (surge, roll) even though heave and pitch spectra may be unaffected.

    \subsection{Frequency Coverage}
    \label{sec:freqcov}

    The RAO grid must span the energy-significant part of $S_\eta$.
    For JONSWAP with $\gamma=3.3$, significant energy exists from roughly
    $0.5\omega_p$ to $3\omega_p$.
    RAO tables truncated near the peak frequency will miss important contributions.

    \subsection{Complex Interpolation}
    \label{sec:interp}

    Interpolating amplitude and wrapped phase separately leads to errors near
    phase jumps of $\pm\pi$.
    Always interpolate the real and imaginary parts of the complex RAO.

    \subsection{Scalar Local Wave Height as Input}
    \label{sec:localeta}

    The vessel should not be driven by a scalar time series sampled from the wave grid at
    one point.
    The RAO already represents the integral effect of the hull geometry on an entire
    incident regular wave.
    The correct inputs are component amplitude, frequency, and heading.

    \subsection{Viscous Roll Limitations}
    \label{sec:roll}

    Roll response near the natural roll period is often strongly affected by viscous
    damping, which is absent from potential-flow panel codes.
    A linear potential-flow roll RAO will generally overpredict roll amplitude near
    resonance unless an empirical damping correction is applied
    (e.g., via Ikeda's method or from model test data).

    \subsection{Linear Range}
    \label{sec:linearrange}

    RAO methods assume linear or weakly nonlinear response.
    They become unreliable for slamming and water impact, green water on deck,
    broaching or surf-riding in following seas, large-amplitude rolling, strongly
    viscous hull--wave interaction, and extreme rogue-wave events.

% ============================================================
    \section{Operability Analysis}
    \label{sec:operability}

    A common engineering deliverable is an operability diagram: a plot of the fraction of
    time a vessel can safely execute a given task as a function of $H_s$ and $T_p$ (or
    vessel heading).

    The standard method is:
    \begin{enumerate}[leftmargin=1.4em]
        \item For each $(H_s,T_p,\psi)$ combination, evaluate response spectra for each
        critical DOF via \eqref{eq:response_spectrum}.
        \item Compute $\sigma_q$ and significant amplitude $q_{1/3}=4\sigma_q$ for each limit
        criterion (e.g., significant heave $<$ 1.5\,m, significant roll $<$ 5°, significant
        crane-tip acceleration $<$ 0.15\,g).
        \item Mark a sea state as \emph{operable} if all criteria are simultaneously satisfied.
        \item Weight by a long-term scatter diagram to compute an overall operability
        percentage.
    \end{enumerate}
    Because \eqref{eq:response_spectrum} is analytic and inexpensive to evaluate, it is
    practical to sweep a fine grid of $(H_s,T_p,\psi)$ values without generating explicit
    time-domain realizations.

% ============================================================
    \section{Compact Mathematical Summary}
    \label{sec:summary}

    Let the existing 3D wave realization be
    \begin{equation}
        \eta(\vect{r},t)
        = \sum_{m=1}^{M}
        a_m\cos\!\left(\vect{k}_m\cdot\vect{r} - \omega_m t + \phi_m\right).
        \label{eq:summary_eta}
    \end{equation}

    \noindent\textbf{Time-domain response} (any linear quantity $q$):
    \begin{equation}
        q(t)
        = \Real\!\left\{
        \sum_{m=1}^{M}
        H_q(\omega_{e,m},\beta_m)\,a_m\,
        \ee^{\ii(\vect{k}_m\cdot\vect{r}_0(t) - \omega_m t + \phi_m)}
        \right\}.
        \label{eq:summary_td}
    \end{equation}

    \noindent\textbf{Response spectrum} (RMS and statistical analysis):
    \begin{equation}
        S_q(\omega)
        = \int_{-\pi}^{\pi}
        |H_q(\omega_e,\beta)|^2\,S_\eta(\omega,\mu)\dd\mu,
        \label{eq:summary_spec}
    \end{equation}
    \begin{equation}
        \sigma_q = \sqrt{\int_0^\infty S_q(\omega)\dd\omega}.
        \label{eq:summary_rms}
    \end{equation}

    \noindent\textbf{Time-domain potential-flow dynamics}
    (mooring, control, transients):
    \begin{multline}
        \bigl(\vect{M}+\vect{A}_\infty\bigr)\ddot{\vect{\xi}}(t)
        + \int_0^t \vect{K}_r(t-\tau)\dot{\vect{\xi}}(\tau)\dd\tau \\
        + \vect{C}_h\vect{\xi}(t)
        = \vect{F}_{\mathrm{exc}}(t) + \vect{F}_{\mathrm{ext}}(t),
        \label{eq:summary_cummins}
    \end{multline}
    with $\vect{F}_{\mathrm{exc}}$ from \eqref{eq:excitation} and the convolution
    optionally replaced by the state-space realization of \cref{sec:statespace}.

    In one sentence:
	\begin{equation}
	\boxed{
	\begin{aligned}
	\text{PM/JONSWAP sea}
	\;&+\;
	\text{vessel hydrodynamic operator}
	\\
	&=\;
	\text{vessel behavior}
	\end{aligned}
	}
	\label{eq:oneliner}
	\end{equation}

% ============================================================
    \section{Scope and Limitations of Linear RAO Methods}
    \label{sec:limits}

    \Cref{tab:scope} summarizes when RAO-based methods are appropriate and when
    higher-fidelity approaches are needed.

    \begin{table}[t]
        \centering
        \caption{Scope and Limitations of Linear Seakeeping Methods}
        \label{tab:scope}
        \small
        \begin{tabular}{@{}L{2.1cm}L{2.5cm}@{}}
            \toprule
            RAOs adequate & Higher fidelity required \\
            \midrule
            Ordinary seakeeping (RMS, spectra) &
            Slamming, green water, deck wetness \\
            Operability analysis &
            Broaching, surf-riding \\
            Irregular-wave motion time series &
            Large-amplitude rolling \\
            Acceleration and habitability &
            Strong viscous coupling \\
            Mooring fatigue (linear load) &
            Rogue-wave impact \\
            Preliminary load estimation &
            CFD-resolved hull--wave interaction \\
            \bottomrule
        \end{tabular}
    \end{table}

% ============================================================
    \section{Conclusion}
    \label{sec:conclusion}

    A 3D PM or JONSWAP free-surface simulation already contains, in its directional
    component list, the complete information required to drive a linear vessel seakeeping
    model.
    The missing ingredient is the vessel hydrodynamic transfer model.
    In linear seakeeping, this is most compactly expressed through complex RAOs: each
    wave component is independently amplitude-scaled and phase-shifted by the RAO, and the
    results are superposed to recover any motion, acceleration, or load time series that
    remains phase-consistent with the original irregular sea realization.
    If only statistical measures are required, the equivalent spectral multiplication
    \eqref{eq:summary_spec} is more efficient and directly yields RMS values, significant
    amplitudes, and spectral moments needed for operability and fatigue analysis.
    When mooring coupling, control feedback, or transient dynamics must be captured, the
    same wave components drive the time-domain Cummins equation \eqref{eq:summary_cummins},
    which can be made computationally tractable through state-space realization of the
    radiation memory kernel.

    The practical bridge is always \eqref{eq:oneliner}.
    For ordinary seakeeping the fastest route is spectral RAO multiplication;
    for phase-consistent time histories it is component-wise RAO synthesis;
    and for fully coupled time-domain dynamics it is the Cummins equation with the
    excitation forces assembled from the same directional wave components.

% ============================================================
    \begin{thebibliography}{99}

        \bibitem{Newman}
        J.~N.~Newman,
        \emph{Marine Hydrodynamics}.
        Cambridge, MA, USA: MIT Press, 1977.

        \bibitem{Faltinsen}
        O.~M.~Faltinsen,
        \emph{Sea Loads on Ships and Offshore Structures}.
        Cambridge, U.K.: Cambridge Univ.\ Press, 1990.

        \bibitem{Perez}
        T.~Perez,
        \emph{Ship Motion Control: Course Keeping and Roll Stabilisation Using Rudder and
        Fins}.
        London, U.K.: Springer, 2005.

        \bibitem{Cummins}
        W.~E.~Cummins,
        ``The impulse response function and ship motions,''
        \emph{Schiffstechnik}, vol.~9, pp.~101--109, 1962.

        \bibitem{Wheeler}
        J.~D.~Wheeler,
        ``Method for calculating forces produced by irregular waves,''
        \emph{J.\ Petrol.\ Technol.}, vol.~22, no.~3, pp.~359--367, Mar.\ 1970.

        \bibitem{Morison}
        J.~R.~Morison, J.~W.~Johnson, and S.~A.~Schaaf,
        ``The force exerted by surface waves on piles,''
        \emph{J.\ Petrol.\ Technol.}, vol.~2, no.~5, pp.~149--154, 1950.

        \bibitem{WAMIT}
        WAMIT, Inc.,
        \emph{WAMIT User Manual, Version 7}.
        Chestnut Hill, MA, USA: WAMIT, Inc., 2016. [Online].
        Available: \url{https://www.wamit.com}

        \bibitem{Capytaine}
        F.~Ancellin and F.~Dias,
        ``Capytaine: a Python-based linear potential flow solver,''
        \emph{J.\ Open Source Softw.}, vol.~4, no.~36, p.~1341, 2019.

        \bibitem{HydroDyn}
        OpenFAST Team,
        \emph{HydroDyn Theory Manual and User Guide}.
        Golden, CO, USA: National Renewable Energy Laboratory, 2021. [Online].
        Available: \url{https://openfast.readthedocs.io}

        \bibitem{MSS}
        T.~I.~Fossen and T.~Perez,
        ``MSS Toolbox: Marine Systems Simulator.''
        Norwegian Univ.\ of Science and Technology, 2004. [Online].
        Available: \url{https://github.com/cybergalactic/MSS}

        \bibitem{DNVGL}
        DNV GL,
        \emph{DNV-RP-C205: Environmental Conditions and Environmental Loads}, ed.\ 2019.
        H{\o}vik, Norway: DNV, 2019.

    \end{thebibliography}

\end{document}
